### 2025-07-29 player-evaluation.qmd
20250729 - tweaks

Many aspects of a player's athletic profile, including tests of speed (40-yard dash), strength (bench press), agility (20-yard shuttle run; three cone drill), and jumping ability (vertical jump; broad jump) are available from the National Football League (NFL) Combine, which is especially relevant for evaluating rookies.
We demonstrate how to import data from the NFL Combine in @sec-combine.
There are also calculators that integrate information about body shape and information from the NFL Combine to determine a player's relative athletic score (RAS) for their position: <https://ras.football/ras-calculator/>

@@

- Is the player on a new team?
- Do they have a new head coach or coordinator?
- How will the team's offensive scheme change?
- Does the team have a better offensive line?
- Does the team have better receiving targets?
- Will the player's position change on the depth chart?
- Will there be more competition for targets?

@@

In terms of consistency, how consistent was the player they from game to game and from season to season?
For instance, we could examine the standard deviations of players' fantasy points across games in a given season.
However, the standard deviation tends to be upwardly biased as the mean increases.
So, we can account for the player's mean fantasy points per game by dividing their game-to-game standard deviation of fantasy points ($s_x$ or $\sigma_x$) by their mean fantasy points across games ($\bar{x}$ or $\mu_x$).
This is known as the coefficient of variation (CV), which is provided in @eq-coefficientOfVariation.

@@

A player's knowledge, intelligence, and reaction time can help them gain an upper-hand even when they may not be the fastest or strongest.
A player's work ethic and mental toughness may help them be resilient and persevere in the face of challenges.
Contact-related incentives may lead a player to put forth greater effort.
For instance, a contract may have a performance incentive that provides a player greater compensation if they achieve a particular performance milestone (e.g., receiving yards).
Another potential incentive is if a player is in what is called their "contract year" (i.e., the last year of their current contract).
If a player is in the last year of their current contract, they have an incentive to perform well so they can get re-signed to a new contract.

>>>>

Many aspects of a player's athletic profile, including tests of speed (40-yard dash), strength (bench press), agility (20-yard shuttle run; three cone drill), and jumping ability (vertical jump; broad jump) are available from the National Football League (NFL) Combine, which is especially relevant for evaluating rookies.
We demonstrate how to import data from the NFL Combine in @sec-combine.
There are also calculators that integrate information about body shape and information from the NFL Combine to determine a player's relative athletic score (RAS) for their position: <https://ras.football/ras-calculator/> and <https://www.nfl.com/combine/iq/>

@@

- Is the player on a new team?
- Do they have a new head coach or coordinator?
- How will the team's offensive scheme change?
- Does the team have a better offensive line?
- Does the team have better receiving targets?
- Will the player's position change on the depth chart?
- Will there be more competition for targets?
- Is there a greater level of competition by opposing teams?

@@

In terms of consistency, how consistent was the player from game to game and from season to season?
For instance, we could examine the standard deviations of players' fantasy points across games in a given season.
However, the standard deviation tends to be upwardly biased as the mean increases.
So, we can account for the player's mean fantasy points per game by dividing their game-to-game standard deviation of fantasy points ($s_x$ or $\sigma_x$) by their mean fantasy points across games ($\bar{x}$ or $\mu_x$).
This is known as the coefficient of variation (CV), which is provided in @eq-coefficientOfVariation.

@@

A player's knowledge, intelligence, and reaction time can help them gain an upper-hand even when they may not be the fastest or strongest.
The Wonderlic test was formerly used to assess a player's cognitive abilities [@OConnell2022; archived at <https://web.archive.org/web/20220303015018/https://www.nytimes.com/2022/03/02/sports/football/nfl-wonderlic-test.html>].
A player's work ethic and mental toughness may help them be resilient and persevere in the face of challenges.
Contact-related incentives may lead a player to put forth greater effort.
For instance, a contract may have a performance incentive that provides a player greater compensation if they achieve a particular performance milestone (e.g., receiving yards).
Another potential incentive is if a player is in what is called their "contract year" (i.e., the last year of their current contract).
If a player is in the last year of their current contract, they have an incentive to perform well so they can get re-signed to a new contract.
